Ecosystem size and complexity dictate riverine biodiversity

### 2021-02-14 manuscript.Rmd
Update manuscript.Rmd

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota]
- Taku Kadoya^[National Institute for Environmental Studies]
- Yusuke Miyazaki^[Shiraume Gakuen University]
output:
  word_document: default
  html_notebook: default
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous pattern[@lomolinoEcologyMostGeneral2000] and has formed the basis of area-based conservation strategies[@desmetUsingSpeciesareaRelationship2004; @neigelSpeciesareaRelationshipsMarine2003]. However, many ecosystems possess a complex spatial structure that cannot be represented by area (the scale-invariant complexity)[@rodriguez-iturbeFractalRiverBasins2001; @turnerLandscapeEcologyTheory2015], and its role in scaling biodiversity is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbellUnifiedNeutralTheory2001; @kallimanisHowDoesHabitat2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of spatial difference in species composition was apparent when simulated species dispersal is limited. Cross-watershed comparisons across two geographic regions (the Hokkaido island in Japan and the midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.
bibliography: reference.bib
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

@@

# Data and code availability
Third parties provided data for empirical data analysis; references are provided in the main text and Supplementary Information where appropriate. R functions for the generation of branching networks and the metacommunity simulation are provided as the R package 'mcbrnet.' (available at https://github.com/aterui/mcbrnet). Codes for simulations, statistical analysis, figures, and tables are available at https://github.com/aterui/public-proj_stream-diversity.

>>>>

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota, St. Paul, MN 55108, USA]
- Taku Kadoya^[National Institute for Environmental Studies, Tsukuba, Ibaraki 305-8506, Japan]
- Yusuke Miyazaki^[Shiraume Gakuen College, Kodaira, Tokyo 187-8570, Japan]
output:
  word_document: default
  html_notebook: default
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous pattern in ecology[@lomolinoEcologyMostGeneral2000] and has formed the basis of area-based conservation strategies[@desmetUsingSpeciesareaRelationship2004; @neigelSpeciesareaRelationshipsMarine2003]. However, many ecosystems possess a complex spatial structure that cannot be represented by area (the scale-invariant complexity)[@rodriguez-iturbeFractalRiverBasins2001; @turnerLandscapeEcologyTheory2015], and its role in scaling biodiversity is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodriguez-iturbeFractalRiverBasins2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbellUnifiedNeutralTheory2001; @kallimanisHowDoesHabitat2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of spatial difference in species composition was apparent when simulated species dispersal is limited. Cross-watershed comparisons across two geographic regions (the Hokkaido island in Japan and the midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.
bibliography: reference.bib
csl: nature.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
---

@@

# Data and code availability
Third parties provided data for empirical data analysis; references are provided in the main text and Supplementary Information where appropriate. R functions for the generation of branching networks and the metacommunity simulation are provided as the R package 'mcbrnet' (available at https://github.com/aterui/mcbrnet). Codes for simulations, statistical analysis, figures, and tables are available at XXX.
